\systemEnding: Bootstrapping Domain-Specific Languages for Neurally-Guided Bayesian Program Learning

### 2019-04-07 archive.tex
updated some learning curves

\begin{figure}
  \includegraphics[width = \textwidth]{figures/DeepLogo.pdf}
  \caption{\textbf{Bottom}: Learned graphics programing DSL. Each circle encloses a different learned component; arrows point from a component to other components that use it. \textbf{Top}: 3 tasks, and below them the programs that solve them. Yellow/Blue/Green highlighting indicates which DSL components are used to solve which tasks.}
  \end{figure}

\begin{figure}
  \begin{tikzpicture}
    \renewcommand{\baselinestretch}{0.5}
    \newcommand{\inventionColor}{orange}
    \newcommand{\name}[2]{
      \mcode{{\color{\inventionColor}$f_{#1}$}$($#2$)\,=\,$}
    }
    \newcommand{\noArguments}[1]{
      \mcode{{\color{\inventionColor}$f_{#1}\,=\,$}}
    }

    \newcommand{\horizontalSpacing}{3.8cm}
    \newcommand{\verticalSpacing}{2cm}

%    \draw (0,0) rectangle (10,5);
    \node(t1)[align=left] at (1,-1) {
      \footnotesize +106 769-43$\to $106.769.43\\
      \footnotesize +83 973-831$\to $83.973.831\\
      \footnotesize $f(\text{\mcode{s}}) = $\mcode{(}\subnode{f01}{${\color{\inventionColor}f_0}$}\mcode{  \codechar{.} \codechar{-}}\\
      \footnotesize \hspace{1.25cm}\mcode{(${\color{\inventionColor}f_0}$ \codechar{.} \codechar{ }}\\
      \footnotesize \hspace{1.4cm}\mcode{(cdr s)))}
    };
    \node(t2)[align=left,anchor=center] at ([xshift=\horizontalSpacing]t1.center) {
      \footnotesize Dream Coder$\to$Dr. Dream\\
      \footnotesize Andy Cencici$\to$Dr. Andy\\
      \footnotesize Jan Kotas$\to$Dr. Jan\\
      \footnotesize Nancy Jones$\to$Dr. Nancy\\
      \footnotesize $f(\text{\mcode{s}}) = $\mcode{(${\color{\inventionColor}f_3}$ (${\color{\inventionColor}f_4}$ \codechar{ } s))}
    };
    \node(t3)[align=left] at ([xshift=2cm]t2.east) {
      \footnotesize Dream (Coder)$\to$Coder\\
      \footnotesize 1 (2) 3 4$\to$2\\
      \footnotesize the (dog) ran$\to$ran\\
      \footnotesize +817(33)99$\to$33\\
      \footnotesize $f(\text{\mcode{s}}) = $\mcode{(${\color{\inventionColor}f_6}$ \codechar{(} \codechar{)} s)}
    };

    \draw (-2,-2.5) -- (15,-2.5);
    \node[align=left,rotate=90] at (-1.5,-1.3) {Tasks and\\\\\texttt{Programs}};

    \node[align=left,rotate=90] at (-1.5,-6.3) {Domain Specific Language (DSL)\\\\System discovers subroutines in {\color{\inventionColor}orange}};

    \newcommand{\invention}[4]{
      \node(#1)[inner sep=1,anchor=center,text width=2cm,draw,rounded corners,align=center] at (#2) {
        \footnotesize #3
      };
      \node[anchor=north] at ([yshift=3.5pt]#1.south) {#4};
    }
    \newcommand{\vinvention}[4]{
      \invention{#1}{[yshift=-\verticalSpacing]#2.center}{#3}{#4}
    }
    \newcommand{\hinvention}[4]{
      \invention{#1}{[xshift=\horizontalSpacing]#2.center}{#3}{#4}
    }

    \invention{substitution}{1,-4}{
      ${\color{\inventionColor}f_0}($\mcode{s}$,$\mcode{a}$,$\mcode{b}$) \,=\, $\\
       \mcode{(map ($\lambda$ (x) (if (= x a)\\ b x)) s)}
    }{substitute chars}
    \vinvention{cat}{substitution}{
      ${\color{\inventionColor}f_1}($\mcode{a}$,$\mcode{b}$) \,=\, $\\
      \mcode{(fold ($\lambda$ (x y)\\ (cons x y))\\ a b))}
    }{concat strings}
    \vinvention{looping}{cat}{
       ${\color{\inventionColor}f_2}($\mcode{c}$,$\mcode{p}$,$\mcode{s}$) \,=\, $\\
      \mcode{(unfold l ($\lambda$ (x) (= c (car x))) p ($\lambda$ (z) (cdr z)))}
    }{loop until char}

    \hinvention{constant}{substitution}{
      ${\color{\inventionColor}f_3}($\mcode{s}$) \,=\, $\\
      \mcode{(${\color{\inventionColor}f_1}$ s string)}
    }{prepend constant}
    \vinvention{drop}{constant}{
      ${\color{\inventionColor}f_4}($\mcode{c}$) \,=\, $\\
      \mcode{(${\color{\inventionColor}f_2}$ c ($\lambda$ }\\\mcode{(z) (car z)))}
    }{prefix}
    \vinvention{suffix}{drop}{
      ${\color{\inventionColor}f_5}($\mcode{s}$,$\mcode{c}$) \,=\, $\\
      \mcode{(fold ($\lambda$ (x y) }\\\mcode{(cdr (if (= x c)}\\\mcode{ s y))) s s)}
    }{suffix}

    \hinvention{delimited}{constant}{
      \mcode{${\color{\inventionColor}f_6}($x$,$y$,$s$)\,=\,$}\\
      \mcode{(${\color{\inventionColor}f_4}$ x (${\color{\inventionColor}f_5}$ y c))}
    }{delimited substring}
    \draw[->] (delimited.west) -- (drop.east);
    \draw[->] (delimited.west) -- (suffix.east);

    \vinvention{appendSingle}{delimited}{
      \mcode{${\color{\inventionColor}f_7}($c$,$s$)\,=\,$}\\
      \mcode{(${\color{\inventionColor}f_1}$ (cons c nil) s)}}{append character}

    \vinvention{surround}{appendSingle}{
      \mcode{${\color{\inventionColor}f_8}($s$)\,=\,$}\\
      \mcode{(cons \codechar{(} (${\color{\inventionColor}f_7}$ \codechar{)} s))}}{surround in parens}
    \draw[->] (surround.west) to[in=200,out=160] ([yshift=-0.25cm]appendSingle.west);
    \draw[->] (appendSingle.west) to[in=40,out=120] ([yshift=0.25cm]cat.east);

    \draw[->] (drop.west) -- (looping.east);
    \draw[->] (constant.west) -- (cat.east);

    \draw[->] ([xshift=-0.5cm]t2.south) -- (constant.north);
    \draw[->] ([xshift=0.3cm]t2.south) to[in=30,out=-30] ([xshift=1cm]drop.north);

    \draw[->] ([yshift=0.2cm,xshift=-0.2cm]t1.south) to[in=100,out=200] (substitution.north);
    \draw[->] ([yshift=0.5cm,xshift=-0.5cm]t1.south) to[in=120,out=200] ([xshift=-0.5cm]substitution.north);
    \draw[->] ([xshift=-0.5cm]t3.south) -- (delimited.north);

    \begin{scope}[shift={(0,-15)}]
      \invention{filter}{0,0}{
        \name{0}{p$,$z}\mcode{\\(fold ($\lambda$ (x a)\\ (if (p x) (cons x a) a))\\ nil z)} 
        }{higher-order `filter'}

      \vinvention{three}{filter}{\footnotesize ${\color{\inventionColor}f_1} = $\\\mcode{(+ 1 (+ 1 1))}}{The number `3'}
      \vinvention{prefix}{three}{\name{2}{n$,$z}\\\mcode{(map ($\lambda$ (i) (index i z)) (range (+ 1 n)))}}{First $n$ elements}
      \hinvention{single}{filter}{\name{3}{z$,$l}\\\mcode{(fold ($\lambda$ (x y)\\(cons x y))\\ z l)}}{Append element}
      \vinvention{for}{single}{\footnotesize ${\color{\inventionColor}f_4} = $\mcode{(+ 1 ${\color{\inventionColor}f_1}$)}}{The number `4'}
      \vinvention{Max}{for}{\name{5}{l}\\\mcode{%
          (car (${\color{\inventionColor}f_0}$ ($\lambda$ (x) (nil? (${\color{\inventionColor}f_0}$ l ($\lambda$ (z) (> z x))))) l))}}{max of list}
      \hinvention{largest}{single}{\name{6}{l$,$n}\\\mcode{(${\color{\inventionColor}f_5}$ (${\color{\inventionColor}f_0}$ ($\lambda$ (x) \\(> n (length (${\color{\inventionColor}f_0}$ ($\lambda$ (y) (> x y)) l)))) l))}}{$n^{\text{th}}$ largest element}
      \vinvention{count}{largest}{\name{7}{p$,$l}\\\mcode{(length\\ (${\color{\inventionColor}f_0}$ ($\lambda$ (y) (= x y)) l))}}{count}
      \vinvention{weird}{count}{\name{8}{p$,$l}\mcode{
          (map ($\lambda$ (x) (if (p x) (+ 1 x) 0))))}}
                 {Higher-order incrementer}

      \draw[->] (for.west) -- (three.east);
      \draw[->] ([yshift=-0.0cm]largest.west) -- (Max.east);
      \draw[->] ([yshift=-0.cm]largest.west) to[out=240,in=330] ([yshift=-0.1cm]filter.east);
      \draw[->] (Max.west) -- (filter.east);
      \draw[->] ([yshift=0.1cm]count.west) -- ([yshift=-0.3cm]filter.east);

      \node(t1)[align=left,anchor=center] at (0,3) {
        \mcode{[1 4 5 8]$\to$[5 8]}\\
        \mcode{[9 8 3 8]$\to$[9 8 8]}\\
        \mcode{[3 4 5 6]$\to$[5 6]}\\
        \mcode{[6 5 4 3]$\to$[6 5]}\\
        \footnotesize $f(\text{\mcode{z}}) = $\mcode{(${\color{\inventionColor}f_0}$ ($\lambda$ (x)}\\\hspace{0.5cm}\mcode{ (> x ${\color{\inventionColor}f_4}$)) z)}
      };
      \node(t2)[align=left] at ([xshift=\horizontalSpacing]t1.center) {
        \mcode{[9 2 4]$\to$[4 2 9]}\\
        \mcode{[4 3]$\to$[3 4]}\\
        \mcode{[3 4 5 6]$\to$[6 5 4 3]}\\
        \mcode{[]$\to$[]}\\
        \footnotesize $f(\text{\mcode{z}}) = $\mcode{(fold ($\lambda$ (x a)}\\\hspace{0.25cm}\mcode{ (${\color{\inventionColor}f_3}$ x a)) nil z)}
      };
      \node(t3)[align=left] at ([xshift=\horizontalSpacing]t2.center) {
        \mcode{[9 2 3]$\to$[2 3 9]}\\
        \mcode{[2 1]$\to$[1 2]}\\
        \mcode{[8 7 9 2 5]$\to$[2 5 7 8 9]}\\
        \footnotesize $f(\text{\mcode{z}}) = $\mcode{(map ($\lambda$ (i)}\\\hspace{0.5cm}\mcode{(${\color{\inventionColor}f_6}$ z i)) }\\\hspace{0.5cm}\mcode{(+ 1 (range }\\\hspace{0.5cm}\mcode{(length z)))}
      };

      \draw[->] ([yshift=0.35cm,xshift=-0.3cm]t1.south) to[out=250] (filter.north);
      \draw[->] ([yshift=0.0cm,xshift=0.3cm]t1.south) to[out=350,in=100] (for.west);
      \draw[->] ([xshift=-0.7cm]t2.south) -- (single.north);
      \draw[->] ([xshift=-1.3cm,yshift=0.6cm]t3.south) to[out=240] (largest.north);

    \end{tikzpicture}
\end{figure}

@@

An old dream within AI is a machine that learns and reasons by writing
its own programs.  This vision stretches back to the
1960's~\cite{solomonoff1964formal} and, if fully realized, could bring
us much closer to machines that learn and think like humans.
Computational models of cognition often explain the flexibility and
richness of human thinking in terms of program learning: from everyday
thinking and problem solving (motor program induction as an account of
recognition and generation of handwriting and
speech~\cite{lake2015human}; functional programs as a model of natural
language semantics~\cite{SOMETHING}) to learning problems that unfold
over longer developmental time scales: the child's acquisition of
intuitive theories (of kinship, taxonomy, etc.)~\cite{Ullman2012} and
natural language grammar~\cite{DBLP:journals/cogsr/SchmidK11}, to name
just a few.  An outstanding challenge, however, is to engineer
program-learners that display the same level of domain-generality as
the humans they are meant to model.

Recent program-learning systems developed within the AI and machine
learning community are impressive along many dimensions, authoring
programs for problem domains like drawing
pictures~\cite{spiral,ellis2017learning}, transforming
text~\cite{gulwani2011automating} and numerical
sequences~\cite{balog2016deepcoder}, robot motion
planning~\cite{devlin2017neural}, and reasoning over common sense
knowledge bases~\cite{muggleton2015meta}.  These systems work in
different ways, but typically hinge upon a carefully hand-engineered
Domain Specific Language (DSL).  The DSL restricts the space of
programs to contain the kinds of concepts needed for one specific
domain.  For example, a picture-drawing DSL could include concepts
like circles and spirals, and a DSL for numerical sequences could
include sorting and reversing lists of numbers.  Modern systems also
learn how to efficiently deploy the DSL on new
problems~\cite{devlin2017robustfill,balog2016deepcoder,NGDS}, but --
unlike human learners -- do not discover the underlying system of
concepts needed to navigate the domain.

We contribute a program-induction system that learns the
domain-specific concepts (DSL) while jointly learning how to use those
concepts.  This joint learning problem models two complementary
notions of domain expertise: domain experts have at their disposal
a powerful, yet specialized repertoire of concepts and abstractions
(analogous to the DSL) while also having accurate intuitions about
when and how to use those concepts to solve new problems.
Representative domains,
along with DSLs we learn for them,
are shown in Figure~\ref{exampleDSL}.

We call our system `DreamCoder' because it acquires these two kinds of
expertise through a novel kind of wake/sleep or `dream'
learning~\cite{hinton1995wake}, iterating through a wake cycle --
where it solves problems by writing programs -- and a pair of sleep
cycles, both of which are loosely biologically inspired by actual
sleep.  The first sleep cycle, which we refer to as
\textbf{consolidation}, grows the DSL by replying experiences from
waking and consolidating them into new code abstractions.  This cycle
is inspired by the formation of abstractions during sleep memory
consolidation~\cite{DUDAI201520}.  The second sleep cycle, which we
refer to as \textbf{dreaming}, improves the agents knowledge of how to
write code by training a neural network to help search for
programs. The neural net is trained on replayed experiences as well as
`fantasies', or samples, from the DSL.  These two kinds of dreams are
inspired by the distinct episodic replay and hallucination components
of dream sleep~\cite{fosse2003dreaming}.

\begin{comment}
This dream-learning architecture brings together two lines of prior work,
both of which have been separately influential within
artificial intelligence.
One line of work
considers the problem of learning new concepts, abstractions, or
`options' from
experience~\cite{Dechter:2013:BLV:2540128.2540316,DBLP:conf/icml/LiangJK10,solomonoff1989system,DBLP:journals/corr/abs-1110-5667,stolle2002learning},
while the other line of work considers the problem of learning how to
deploy those concepts
efficiently~\cite{devlin2017robustfill,balog2016deepcoder,NGDS}.  Our
goal with DreamCoder is to show that the combination of these ideas is
uniquely powerful, and pushes us toward program-writing systems that,
like human learners, autonomously acquire the expertise needed to
navigate a new domain of problems.
\end{comment}

Each wake/sleep cycle creates new DSL components that build on
components learned in earlier sleep cycles, growing a DSL with nested 
hierarchies of code.  We identify this cumulative nesting of
abstractions as a variety of deep representation
learning~\cite{lecun2015deep}.  Figure~\ref{exampleDSL} diagrams a
subset of these learned networks (the DSL). For example, the system
learns to sort sequences of numbers by invoking a DSL component 4
layers deep, or draws the leftmost pairs of images in
Figure~\ref{exampleDSL} using a depth-3 component.  For this reason we
refer to \system as an instance of `deep program learning'.

Our goal with \system is to engineer a system that develops domain
expertise in humanlike ways. This involves learning both declarative and
procedural knowledge, like a DSL and a recognition model, but includes
other features of the development of expertise:
humans can become experts in many fields,
and so we evaluate our algorithm across six different problem domains;
a human expert doesn't become an expert overnight,
and needs more than a handful of example problems to learn from,
but doesn't need millions of examples;
similarly our algorithm's
learning trajectory unfolds over a series of wake/sleep cycles
requiring around a hundred problems per domain.

\begin{figure}\centering
  \begin{tikzpicture}
    \renewcommand{\baselinestretch}{0.5}
    \newcommand{\inventionColor}{orange}

    \newcommand{\name}[2]{
      \mcode{{\color{\inventionColor}$f_{#1}$}$($#2$)\,=\,$}
    }
    \newcommand{\noArguments}[1]{
      \mcode{{\color{\inventionColor}$f_{#1}$}$\,=\,$}
    }
    \newcommand{\use}[1]{{\color{\inventionColor}$f_{#1}$}}
    \newcommand{\horizontalSpacing}{4.2cm}
    \newcommand{\verticalSpacing}{2cm}

    \newcommand{\invention}[4]{
      \node(#1)[inner sep=1,anchor=south,text width=2.2cm,draw,rounded corners,align=center] at (#2) {
        \footnotesize #3
      };
      \node[anchor=north,align=center] at ([yshift=3.5pt]#1.south) {#4};
    }
    \newcommand{\vinvention}[4]{
      \invention{#1}{[yshift=-\verticalSpacing]#2.south}{#3}{#4}
    }
    \newcommand{\hinvention}[4]{
      \invention{#1}{[xshift=\horizontalSpacing]#2.south}{#3}{#4}
    }
    \newcommand{\exampleFunction}[2]{
      \includegraphics[width = 0.5cm]{#1}, \includegraphics[width = 0.5cm]{#2}
    }
    \newcommand{\task}[4]{
      \node[align=left,anchor=center](t#1) at (#2) {#3};
      \node[align=left](p#1) at ([yshift=-0.75cm]t#1.south) {#4};
    }
    \newcommand{\nextTask}[4]{
      \task{#1}{[xshift=\horizontalSpacing]#2.center}{#3}{#4}
    }

    \begin{scope}[shift={(0,8)}]
      \task{1}{2,3}{
        \textsc{Reverse}\\
        \mcode{[9 2 4]$\to$[4 2 9]}\\
        \mcode{[4 3]$\to$[3 4]}\\
        \mcode{[3 4 5 6]$\to$[6 5 4 3]}\\
        \mcode{[]$\to$[]}
      }{\footnotesize $f(\text{\mcode{z}}) = $\mcode{(fold ($\lambda$ (x a)}\\\hspace{0.25cm}\mcode{ (${\color{\inventionColor}f_3}$ x a)) nil z)}}
      \nextTask{2}{t1}{
        \textsc{Take if $\geq 5$}\\
        \mcode{[1 4 5 8]$\to$[5 8]}\\
        \mcode{[9 8 3 8]$\to$[9 8 8]}\\
        \mcode{[3 4 5 6]$\to$[5 6]}\\
        \mcode{[6 5 4 3]$\to$[6 5]}
      }{\footnotesize $f(\text{\mcode{z}}) = $\mcode{(${\color{\inventionColor}f_0}$ ($\lambda$ (x)}\\\hspace{0.5cm}\mcode{ (> x ${\color{\inventionColor}f_4}$)) z)}}

      \nextTask{3}{t2}{
        \textsc{Sort}\\
        \mcode{[9 2 3]$\to$[2 3 9]}\\
        \mcode{[2 1]$\to$[1 2]}\\
        \mcode{[8 7 9 2 5]$\to$[2 5 7 8 9]}\\
        \mcode{[9 8 9]$\to$[8 9 9]}
      }{\footnotesize $f(\text{\mcode{z}}) = $\mcode{(map ($\lambda$ (i)}\\\hspace{0.5cm}\mcode{(${\color{\inventionColor}f_6}$ z i)) }\\\hspace{0.5cm}\mcode{(+ 1 (range }\\\hspace{0.5cm}\mcode{(length z)))}}

      %% \vinvention{filter}{p1}{
      %%   \name{0}{p$,$z}\mcode{\\(fold ($\lambda$ (x a)\\ (if (p x) (cons x a) a))\\ nil z)} 
      %% }{higher-order `filter'}
      \vinvention{single}{p1}{\name{3}{z$,$l}\\\mcode{(fold ($\lambda$ (x y)\\(cons x y))\\ z l)}}{Append element}

      \vinvention{three}{single}{\footnotesize \noArguments{1}\\\mcode{(+ 1 (+ 1 1))}}{The number `3'}
      \vinvention{prefix}{three}{\name{2}{n$,$z}\\\mcode{(map ($\lambda$ (i) (index i z)) (range (+ 1 n)))}}{First $n$ elements}
      \hinvention{filter}{single}{
        \name{0}{p$,$z}\mcode{\\(fold ($\lambda$ (x a)\\ (if (p x) (cons x a) a))\\ nil z)} 
      }{higher-order `filter'}
      \vinvention{for}{filter}{\noArguments{4}\\\mcode{(+ 1 ${\color{\inventionColor}f_1}$)}}{The number `4'}
      \vinvention{weird}{for}{\name{8}{p$,$l}\mcode{
          (map ($\lambda$ (x) (if (p x) (+ 1 x) 0)))}}{Higher-order incrementer}                 
      \hinvention{largest}{filter}{\name{6}{l$,$n}\\\mcode{(${\color{\inventionColor}f_5}$ (${\color{\inventionColor}f_0}$ ($\lambda$ (x) \\(> n (length (${\color{\inventionColor}f_0}$ ($\lambda$ (y) (> x y)) l)))) l))}}{$n^{\text{th}}$ largest element}
      \vinvention{Max}{largest}{\name{5}{l}\\\mcode{%
          (car (${\color{\inventionColor}f_0}$ ($\lambda$ (x) (nil? (${\color{\inventionColor}f_0}$ l ($\lambda$ (z) (> z x))))) l))}}{max of list}
      \vinvention{count}{Max}{\name{7}{p$,$l}\\\mcode{(length\\ (${\color{\inventionColor}f_0}$ ($\lambda$ (y) (= x y)) l))}}{count}

      \node[rotate=90,align=center] at ([xshift=-2cm]t1.center) {\textbf{\textsc{Tasks}}\\};
      \node(programLabel)[rotate=90,align=center] at ([xshift=-2cm]p1.center) {\texttt{\textbf{Progs.}}\\};
      \node[rotate=90,align=center] at ([xshift=-2cm]three.center) {\textbf{\textsc{Domain Specific Language}}\\
        \small Learned routines in {\color{orange}orange}};
      \node at ([yshift=0.5cm]t2.north) {\textbf{\Large \textsc{Domain: List Processing}}};

      \draw [dashed] ([xshift=0.5cm,yshift=-0.25\verticalSpacing]programLabel.east) -- ([xshift=15cm,yshift=-0.25\verticalSpacing]programLabel.east);
    \node[align=left,anchor=center] at ([xshift=14cm,yshift=-0.3\verticalSpacing]programLabel.east) {\large $\uparrow$\;system inputs\\\\\\\large $\downarrow$\;system outputs};

      \draw[->] (for.west) -- (three.east);
      \draw[->] (largest.east) to[out=0,in=0] (Max.east);
      \draw[->] ([yshift=-0.cm]largest.west) -- ([yshift=0.2cm]filter.east);
      \draw[->] (Max.west) -- (filter.east);
      \draw[->] ([yshift=0.1cm]count.west) -- ([yshift=-0.3cm]filter.east);

      \draw[->] ([yshift=0.35cm,xshift=-0.3cm]p2.south) to[out=250] (filter.north);
      \draw[->] ([yshift=0.0cm,xshift=0.3cm]p2.south) to[out=350,in=0] (for.east);
      \draw[->] ([xshift=-0.7cm]p1.south) -- (single.north);
      \draw[->] ([xshift=-1.3cm,yshift=0.6cm]p3.south) to[out=240] (largest.north);

    \renewcommand{\task}[4]{
      \node(t#1) at (#2) {\includegraphics[width = 2cm]{#3}};
      \node(p#1) at ([yshift=-0.75cm]t#1.south) {#4};
    }
    \newcommand{\nextTask}[4]{
      \task{#1}{[xshift=\horizontalSpacing]#2.center}{#3}{#4}
    }

    \task{1}{2,0}{figures/logoTasks/logo144.png}{\mcode{(for (i 4) (\use{3} i))}}
    \nextTask{2}{t1}{figures/logoTasks/logo121.png}{\mcode{(\use{7} 6 (\use{1}; (\use{0} 1)))}}
    \nextTask{3}{t2}{figures/logoTasks/logo133.png}{\mcode{(\use{7} 7 (\use{8}; \use{6}))}}

    \vinvention{circle}{p1}{
      \name{0}{r}\mcode{(for (i $\infty$) (move r $\epsilon$) (move $\epsilon$ $\epsilon$))}
    }{circle\\\exampleFunction{figures/logoTasks/logo51}{figures/logoTasks/logo63_dilated}}
    \vinvention{pu}{circle}{
      \noArguments{1}\mcode{(pen-up (move 1 0))}
    }{pick up pen\\and move}
    \vinvention{repeated}{pu}{
      \name{2}{n$,$d$,$a}\mcode{(for (i n) (move d a))}
    }{repeated segment\\\exampleFunction{figures/logoTasks/logoDemo1_dilated}{figures/logoTasks/logoDemo0_dilated.png}}
    \hinvention{square}{circle}{
      \name{3}{d}\mcode{(\use{2} 4 d (/ $2\pi$ 4))}
    }{square\\\exampleFunction{figures/logoTasks/logo139_dilated}{figures/logoTasks/logo142_dilated}}
    \vinvention{smooth}{square}{
      \name{4}{d$,$a}\mcode{(\use{2} $\infty$ d a)}
    }{Smooth curve\\\exampleFunction{figures/logoTasks/logoDemo2_dilated}{figures/logoTasks/logoDemo3_dilated}}
    \vinvention{angle}{smooth}{
      \name{5}{n}\mcode{(/ $2\pi$ n)}
    }{$\frac{2\pi}{n}$}
    \hinvention{semi}{square}{
      \noArguments{6}\mcode{(\use{4} $\epsilon$ $\epsilon$)}
    }{semicircle\\\exampleFunction{figures/logoTasks/logo55_dilated}{figures/logoTasks/logo62_dilated}}
    \vinvention{snow}{semi}{
      \name{7}{n$,$f}\mcode{(for (i n) \\(move 0 (\use{5} n))\\ (get/set f))}
    }{rotational symmetry\\\exampleFunction{figures/logoTasks/logo119_dilated}{figures/logoTasks/logo68_dilated}}
    \vinvention{line}{snow}{
      \noArguments{8}\mcode{(move 1 0)}
    }{line segment}

    \draw[->] (p1.south) -- (square.north);
    \draw[->] (p2.south) -- (circle.north);
    \draw[->] (p3.south) -- (semi.north);
    \draw[->] (p2.south) -- (snow.west);
    \draw[->] (p2.south) -- (pu.east);
    \draw[->] (p3.south) to[out=190,in=120] (line.west);
    \draw[->] (p3.south) to[out=200,in=120] (snow.west);

>>>>

@@

What would it take to build a machine that can learn everything a
person does over their lifetime?
Although this goal will remain distant for the foreseeable future,
we know that a general learning system like this would need to be able to acquire many different kinds of expertise.
Virtually every child becomes an expert in natural language, motor control, and social interaction, while learning rich
intuitive theories of kinship, taxonomy and physics. Many of those
children will grow into adults with competence in cooking, calculus, tennis, drawing
pictures, or writing software.
Expertise -- the ability to quickly solve new problems in a domain -- is crucial.
Human learning goes far beyond just memorizing a large set of facts or routines,
but involves higher-order skills,
like learning to learn new concepts or
learning to solve new types of problems more quickly and effectively.
Despite great advances in machine learning,
we are still far from
an AI with these abilities.
The modern AI toolkit has given us machines that learn to play challenging games at superhuman levels,
but cannot quickly transfer to similar games like a person would;
or, which can learn to generate convincing English prose,
but which do not then learn to analyze many different languages,
like a field linguist would.
Here, we contribute a computational model that takes a small step toward the goal of building machines that
grow into domain experts,
looking at much simpler kinds of problems that humans can learn to solve and indeed become experts in.

Our model is structured
around the hypothesis that
domain expertise
consists primarily of two ingredients.
First, domain experts have 
explicit, declarative concepts that are powerful yet finely-tuned.
A visual artist has concepts like arcs, spirals, symmetries, and perspectives;
a physicist has concepts like dot products, vector fields, and conservation laws;
and an architect has concepts like arches, supports, and bridges.
Second, experts have implicit procedural skill in deploying those concepts
to quickly solve new problems:
at a glance, human domain
experts can intuit which compositions of concepts are likely to solve
the task at hand, even before searching for a solution.
Studies of human expertise
find that,
compared to novices,
experts can quickly categorize and classify problems based on the ``deep structure''
of the problem's solution~\cite{chi1981expertise,chi2012seeing}.
In short, we take the stance that expertise means both
having the right concepts, and being able to quickly see how those concepts can be composed into a solution.

Because solutions to many kinds of
problems can often be described as some kind of
program~\cite{lake2015human,Ullman2012,DBLP:journals/cogsr/SchmidK11},
our model solves a specific problem by synthesizing a program solving it. It
gradually grows its domain-specific knowledge for a class of problems
by assembling a library of code containing concepts useful for the
domain, and by training a neural network to quickly infer, for a
specific problem, what code in its library is likely to
solve it. We think of these two learned components as analogous
to the explicit concepts and
implicit procedural skill that human experts develop.
We will first investigate our model within classic program
synthesis domains for manipulating sequences of numbers and text, and
then consider visual and creative programs for drawing pictures and
building towers out of toy blocks, and finally consider programs for basic kinds of equation discovery and programming language design, altogether considering both
deterministic and probabilistic programs that act both generatively
(e.g., producing an artifact like an image or plan) and conditionally
(e.g., mapping inputs to outputs).

Our model iteratively creates new library routines that build on
concepts learned earlier in its learning trajectory, growing a library with nested
hierarchies of code.  We think of this cumulative nesting of
abstractions as a variety of deep representation
learning~\cite{lecun2015deep}.  Figure~\ref{exampleDSL} diagrams a
subset of these learned networks (the library). For example, the model
learns to sort sequences of numbers by invoking a library component 4
layers deep, or draws the leftmost pairs of images in
Figure~\ref{exampleDSL} using a depth-3 component.  For this reason we
refer to our approach as an instance of `deep program learning'.

\begin{figure}
  \includegraphics[width = \textwidth]{figures/dpl/list}
  \includegraphics[width = \textwidth]{figures/dpl/logo}
  \caption{}\label{exampleDSL}
\end{figure}

%% \section{Old introduction}

%% An old dream within AI is a machine that learns and reasons by writing
%% its own programs.  This vision stretches back to the
%% 1960's~\cite{solomonoff1964formal} and, if fully realized, could bring
%% us much closer to machines that learn and think like humans.
%% Computational models of cognition often explain the flexibility and
%% richness of human thinking in terms of program learning: from everyday
%% thinking and problem solving (motor program induction as an account of
%% recognition and generation of handwriting and
%% speech~\cite{lake2015human}; functional programs as a model of natural
%% language semantics~\cite{SOMETHING}) to learning problems that unfold
%% over longer developmental time scales: the child's acquisition of
%% intuitive theories (of kinship, taxonomy, etc.)~\cite{Ullman2012} and
%% natural language grammar~\cite{DBLP:journals/cogsr/SchmidK11}, to name
%% just a few.  An outstanding challenge, however, is to engineer
%% program-learners that display the same level of domain-generality as
%% the humans they are meant to model.

%% Recent program-learning systems developed within the AI and machine
%% learning community are impressive along many dimensions, authoring
%% programs for problem domains like drawing
%% pictures~\cite{spiral,ellis2017learning}, transforming
%% text~\cite{gulwani2011automating} and numerical
%% sequences~\cite{balog2016deepcoder}, robot motion
%% planning~\cite{devlin2017neural}, and reasoning over common sense
%% knowledge bases~\cite{muggleton2015meta}.  These systems work in
%% different ways, but typically hinge upon a carefully hand-engineered
%% Domain Specific Language (DSL).  The DSL restricts the space of
%% programs to contain the kinds of concepts needed for one specific
%% domain.  For example, a picture-drawing DSL could include concepts
%% like circles and spirals, and a DSL for numerical sequences could
%% include sorting and reversing lists of numbers.  Modern systems also
%% learn how to efficiently deploy the DSL on new
%% problems~\cite{devlin2017robustfill,balog2016deepcoder,NGDS}, but --
%% unlike human learners -- do not discover the underlying system of
%% concepts needed to navigate the domain.

%% We contribute a program-induction system that learns the
%% domain-specific concepts (DSL) while jointly learning how to use those
%% concepts.  This joint learning problem models two complementary
%% notions of domain expertise: domain experts have at their disposal
%% a powerful, yet specialized repertoire of concepts and abstractions
%% (analogous to the DSL) while also having accurate intuitions about
%% when and how to use those concepts to solve new problems.
%% Representative domains,
%% along with DSLs we learn for them,
%% are shown in Figure~\ref{exampleDSL}.

%% We call our system `DreamCoder' because it acquires these two kinds of
%% expertise through a novel kind of wake/sleep or `dream'
%% learning~\cite{hinton1995wake}, iterating through a wake cycle --
%% where it solves problems by writing programs -- and a pair of sleep
%% cycles, both of which are loosely biologically inspired by actual
%% sleep.  The first sleep cycle, which we refer to as
%% \textbf{consolidation}, grows the DSL by replying experiences from
%% waking and consolidating them into new code abstractions.  This cycle
%% is inspired by the formation of abstractions during sleep memory
%% consolidation~\cite{DUDAI201520}.  The second sleep cycle, which we
%% refer to as \textbf{dreaming}, improves the agents knowledge of how to
%% write code by training a neural network to help search for
%% programs. The neural net is trained on replayed experiences as well as
%% `fantasies', or samples, from the DSL.  These two kinds of dreams are
%% inspired by the distinct episodic replay and hallucination components
%% of dream sleep~\cite{fosse2003dreaming}.

%% This dream-learning architecture brings together two lines of prior work,
%% both of which have been separately influential within
%% artificial intelligence.
%% One line of work
%% considers the problem of learning new concepts, abstractions, or
%% `options' from
%% experience~\cite{Dechter:2013:BLV:2540128.2540316,DBLP:conf/icml/LiangJK10,solomonoff1989system,DBLP:journals/corr/abs-1110-5667,stolle2002learning},
%% while the other line of work considers the problem of learning how to
%% deploy those concepts
%% efficiently~\cite{devlin2017robustfill,balog2016deepcoder,NGDS}.  Our
%% goal with DreamCoder is to show that the combination of these ideas is
%% uniquely powerful, and pushes us toward program-writing systems that,
%% like human learners, autonomously acquire the expertise needed to
%% navigate a new domain of problems.

%% Each wake/sleep cycle creates new DSL components that build on
%% components learned in earlier sleep cycles, growing a DSL with nested 
%% hierarchies of code.  We identify this cumulative nesting of
%% abstractions as a variety of deep representation
%% learning~\cite{lecun2015deep}.  Figure~\ref{exampleDSL} diagrams a
%% subset of these learned networks (the DSL). For example, the system
%% learns to sort sequences of numbers by invoking a DSL component 4
%% layers deep, or draws the leftmost pairs of images in
%% Figure~\ref{exampleDSL} using a depth-3 component.  For this reason we
%% refer to \system as an instance of `deep program learning'.

%% Our goal with \system is to engineer a system that develops domain
%% expertise in humanlike ways. This involves learning both declarative and
%% procedural knowledge, like a DSL and a recognition model, but includes
%% other features of the development of expertise:
%% humans can become experts in many fields,
%% and so we evaluate our algorithm across six different problem domains;
%% a human expert doesn't become an expert overnight,
%% and needs more than a handful of example problems to learn from,
%% but doesn't need millions of examples;
%% similarly our algorithm's
%% learning trajectory unfolds over a series of wake/sleep cycles
%% requiring around a hundred problems per domain.

%\input{dpl1.tex} % old latex deep programming figures
